On Sample Size Needed for Block Bootstrap Confidence Intervals to Have Desired Coverage Rates

### 2023-12-25 Manuscript/reply.tex
First Draft of Replies to Referee 2

@@

Here is an example of what this would look like. We find that it is a little
cluttered to make comparisons, so we will use the original format.

@@

We have added this comment to the discussion:
Lastly, because of its exceptional performance in comparison to the other 
CIs that are not based on bootstrap "tables", the proposed recentered percentile 
CI would be
an interesting method to study further with different bootstrap schemes.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/alt2plot_norm_phi_1}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for 
    the first-order autocorrelation coefficient $\phi$ of an AR(1) process with 
    a marginal standard normal distribution with 
    $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates of
    block bootstrap with $l = \lceil n^{1/3} \rceil$. The
    error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:phi1}
\end{figure}

@@

- When there is a positive correlation, effective sample size is lesser
- When there is a negative correlation, effective sample size is larger, 
meaning the coverage is better

@@

The original sentence has been replaced with:
Coverage rates are acceptable at smaller sample 
sizes when phi is positive versus when phi is negative. In
other words, a larger sample size is generally required to estimate a 
parameter for a sample with a negative $\phi$ versus a positive $\phi$ of the 
same magnitude.

@@

We have now repeated the same simulation study with block size as an 
additional experimental factor. The optimal order of block size is 
$\lceil n^{1/3} \rceil$, so we tried the same simulations using 
$l = \lceil n^{1/3} \rceil$ and $l = \lceil 2n^{1/3} \rceil$

@@

For practical purposes, it is in fact rare for autocorrelations to be stronger
than 0.4. We have established the trend as the strength of the autocorrelation 
increases, and how it varies depending on the sign of the autocorrelation and 
the parameter of interest.

>>>>

Below is an example of what this would look like. We find that it is a little
cluttered to make comparisons, so we will use the original format.

@@

@@

We have added this comment to the discussion:

Lastly, because of its exceptional performance in comparison to the other 
CIs that are not based on bootstrap "tables", the proposed recentered percentile 
CI would be
an interesting method to study further with different bootstrap schemes.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/alt2plot_norm_phi_1}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for 
    the first-order autocorrelation coefficient $\phi$ of an AR(1) process with 
    a marginal standard normal distribution with 
    $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates of
    block bootstrap with $l = \lceil n^{1/3} \rceil$. The
    error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:alt_phi1}
\end{figure}

@@

If a stationary series has a positive autocorrelation, the effective sample
size is decreased, whereas is a series has a negative autocorrelation, the
effective sample size is increased. Additionally, we expect this to have a 
greater effect on the the estimation of the location parameter versus that
of the scale parameter or temporal dependence parameter.

@@

The original sentence has been replaced with:

Coverage rates are acceptable at smaller sample 
sizes when phi is positive versus when phi is negative. In
other words, a larger sample size is generally required to estimate a 
parameter for a sample with a negative $\phi$ versus a positive $\phi$ of the 
same magnitude.

@@

We have now repeated the same simulation study with block size as an 
additional experimental factor. The optimal order of block size is 
$\lceil n^{1/3} \rceil$, so we tried the same simulations using 
$l = \lceil n^{1/3} \rceil$ and $l = \lceil 2n^{1/3} \rceil$. We have included
the changed Results subsection below:

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_norm_mu_1}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the marginal mean $\mu$ of an AR(1) process with a marginal standard 
    normal distribution with AR coefficient
    $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates of
    block bootstrap with $l = \lceil n^{1/3} \rceil$. The
    error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:mu1}
\end{figure}

\begin{figure}[bp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_norm_sigma_1}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the marginal standard deviation $\sigma_x$ of an AR(1) process with a
    marginal standard normal distribution with AR 
    coefficient $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length 
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates of
    block bootstrap with $l = \lceil n^{1/3} \rceil$. The 
    error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:sigma1}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_norm_phi_1}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for 
    the first-order autocorrelation coefficient $\phi$ of an AR(1) process with 
    a marginal standard normal distribution with 
    $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates of
    block bootstrap with $l = \lceil n^{1/3} \rceil$. The
    error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:phi1}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_norm_mu_2}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the marginal mean $\mu$ of an AR(1) process with a marginal standard 
    normal distribution with AR coefficient
    $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates of
    block bootstrap with $l = \lceil 2n^{1/3} \rceil$. The
    error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:mu2}
\end{figure}

\begin{figure}[bp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_norm_sigma_2}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the marginal standard deviation $\sigma_x$ of an AR(1) process with a
    marginal standard normal distribution with AR 
    coefficient $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length 
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates of
    block bootstrap with $l = \lceil 2n^{1/3} \rceil$. The 
    error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:sigma2}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_norm_phi_2}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for 
    the first-order autocorrelation coefficient $\phi$ of an AR(1) process with 
    a marginal standard normal distribution with 
    $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ and series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates of
    block bootstrap with $l = \lceil 2n^{1/3} \rceil$. The
    error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:phi2}
\end{figure}

For estimating the mean parameter~$\mu$, \textbf{Figure~\ref{fig:mu1}} suggests 
that all methods eventually approach correct coverage of $\mu$ as sample size 
increases. Student's $t$ CIs appear to need the smallest sample size to achieve 
correct coverage, except for samples with strong negative dependence, in which 
case, they actually over-cover $\mu$ for smaller sample sizes. For instance, for 
a sample with $n = 100$ and $\phi = -0.4$, the lower bound for a Students $t$ 
CI's coverage of $\mu$ is greater than 95\%, whereas the coverage intervals for 
other methods contain 95\%. The standard normal, percentile, BC, and BCA, and 
recentered percentile CIs require similar sample sizes to recover $\mu$ at the 
nominal level for all combinations of $n$ and $\phi$. All methods seem to 
require a smaller sample to recover $\mu$ at the nominal rate when dealing with 
negative dependence versus positive dependence. For example, BC CIs recover 
$\mu$ for $n \geq 100$ when $\phi = -0.2$, but they only recover $\mu$ for 
$n \geq 800$ when $\phi = 0.2$. In addition, as a negative dependence gets 
stronger, holding everything else equal, coverage increases, which lead to the 
Student $t$ CI's aforementioned over-coverage. As a positive dependence gets 
stronger, holding everything else equal, coverage decreases, and a larger sample 
is necessary to recover~$\mu$.

For estimating the standard deviation parameter~$\sigma_x$,
\textbf{Figure~\ref{fig:sigma1}} suggests that every method can reach nominal 
coverage of $\sigma_x$ if the sample is large enough, but for a given $n$ and 
$\phi$, coverage of $\sigma_x$ will be lower than coverage of $\mu$ in general.
Like $\mu$, $\sigma_x$ can be covered by Student $t$ CIs with smaller sample 
sizes when compared to other methods. Unlike $\mu$, there is no over-coverage 
issue for $\sigma_x$ when $\phi = -0.4$. Standard normal, percentile, BC, BCA, and 
recentered percentile CIs again have similar performance. All methods seem to 
have slightly higher coverage of $\sigma_x$ when $\phi$ is negative versus when 
$\phi$ is positive. Regardless of the sign, coverage of $\sigma_x$ gets worse as 
the strength of the temporal dependence increases.

For estimating the autocorrelation parameter~$\phi$, 
\textbf{Figure~\ref{fig:phi1}} suggests that while standard normal, Student's $t$, and 
recentered percentile CIs do approach correct coverage as sample size increases, 
percentile, BC, and BCA CIs deteriorate as sample size increases, especially as 
the strength of the temporal dependence increases. Because of this, only 
standard normal, Student's $t$, and recentered percentile CIs should be considered as 
effective block bootstrap methods to estimate $\phi$. Student's $t$ CIs once 
again can achieve correct coverage with smaller sample sizes when compared to 
standard and recentered percentile CIs which perform similarly. Student's $t$ 
CIs can can recover $\phi$ at the nominal level for $n \geq 100$ when the 
sample's temporal dependence is as strong as 0.4. Coverage appears to be higher 
for all methods when the dependence is negative rather than positive. Whether or 
not the dependence is negative or positive, coverage of $\phi$ seems to increase 
slightly as the absolute value increases for standard normal, Student's $t$, and 
recentered percentile CIs. For the values of $\phi$ observed, there are no 
examples of over-coverage for standard normal, Student's $t$, BC, BCA, or recentered 
percentile CIs. However, percentile CIs appear to over-cover $\phi$ for smaller 
sample sizes when $\phi = -0.2$ and when $\phi = 0$, indicating again that they 
should not be used.

Based on \textbf{Figure~\ref{fig:mu2}}, for negative autocorrelations, the 
coverage 
rates of $\mu$ appear to be lower when
using $l = \lceil 2n^{1/3} \rceil$. For example, whereas $n = 100$ or 200 
would seem sufficient for most CIs when using $l = \lceil n^{1/3} \rceil$, 
$n = 800$ or 1600 is necessary to capture negative autocorrelations for 
$l = \lceil n^{1/3} \rceil$. Student's $t$ CIs do not seem to be as affected
by this change in $l$: for $\phi = -0.4$ and -0.2, they still overcover $\mu$
for smaller values of $n$.

\textbf{Figure~\ref{fig:sigma2}} looks very similar to 
\textbf{Figure~\ref{fig:sigma1}}, but coverage rates of $\sigma$ do look 
slightly lower
especially for negative values of $\phi$, although Student's $t$ CIs are again
not as influenced by this change in $l$. A larger sample size seems necessary
when using other CIs to estimate $\sigma$ for $l = \lceil 2n^{1/3} \rceil$.

As shown in \textbf{Figure~\ref{fig:phi2}},
recentered percentile and standard CIs have slightly lower coverage rates when
estimating negative values of $\phi$ with $l = \lceil 2n^{1/3} \rceil$. 
Although it is still a problem, the coverage deterioration appears to be less 
dramatic for BCA,
BC, and percentile CIs.

Aside from these differences, the overall changes
in performance when other experimental factors are changed are the same.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/norm_phi_intervals}
  \caption{50 replicate percentile, BC, and recentered percentile CIs for 
    samples of $n \in \{800, 1600\}$ ($l = \lceil n^{1/3} \rceil$)
    for an AR(1) process with $\phi = 0.4$. For 
    each replicate, the lower and upper bounds of the CIs are displayed, as well 
    as $\hat\theta_n$ (blue circle) and $\bar\theta_n^{B}$ (red cross).}
  \label{fig:npi}
\end{figure}

The outcomes of the $\phi$ estimation raise a natural question about the 
lackluster performance of certain methodologies. To delve into this inquiry, a
set of 50 CIs was generated for each of the percentile, BC, and recentered
percentile  approaches for samples of $n \in \{800, 1600\}$. Illustrated in 
\textbf{Figure~\ref{fig:npi}}, it becomes evident that the percentile-based CIs 
exhibit a notable bias, predominantly manifesting as a substantial 
underestimation of $\phi$ with point estimator $\bar\theta_n^B$, that is, the 
average of $B$ bootstrap point estimates. As the sample size increases from 800 
to 1600, this bias does not vanish while the uncertainty reduces, which explains 
why the coverage rates deteriorate. The bias in $\bar\theta_n^B$ also appears to 
invalidate the bias-correction in the BC bootstrap, leading to the poor 
performance of the BC intervals. The BCA intervals have the same problem as the 
BC intervals in the bias-correction step. The root of the issue appears to be 
that the autocorrelation in the block bootstrap samples is somehow smaller 
compared to that in the original sample. On the other hand, the original point 
estimator $\hat\theta_n$ is asymptotically unbiased. Since the width is based on 
the uncertainty in the bootstrap point estimates $\hat\theta_n^{(b)}$, 
$b = 1, \ldots, B$, the percentile CIs recentered at the original point estimate
$\hat\theta_n$ provide desired coverage.

To summarize, the performance of the CIs depends on the target parameter. When
estimating $\mu$ and $\sigma_x$, any CI will do, although Student's $t$ CIs
perform noticeably better than the others.
However, when estimating $\phi$, the 
choice of method is of utmost importance as to avoid coverage deterioration. 
In addition, choice of block size is another important parameter, as
$l = \lceil n^{1/3} \rceil$ results in better coverage than 
$l = \lceil 2n^{1/3} \rceil$.
Coverage rates are acceptable at smaller sample 
sizes when phi is positive versus when phi is negative. In
other words, a larger sample size is generally required to estimate a 
parameter for a sample with a negative $\phi$ versus a positive $\phi$ of the 
same magnitude. In order 
to know if coverage will increase as the strength of the temporal dependence 
increases, one need to know what the parameter of interest is, and in the case 
of $\mu$, the direction of the serial dependence. The BC approach does not seem 
to be correcting bias appropriately when estimating $\phi$. Like the percentile 
method, the recentered percentile method uses the spread from the bootstrap to 
construct the width of the CI. However, the recentered approach, does not 
correct from the original point estimate $\hat\theta_n$.

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_exp_mu_1}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the marginal mean $\mu$ of a stationary series with marginal unit exponential
    distribution obtained by transforming an AR(1) process with
    $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ with series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates of
    block bootstrap with $l = \lceil n^{1/3} \rceil$. 
    The error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:exp_mu1}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_exp_sigma_1}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the marginal standard deviation $\sigma_w$ %\eds{$\sigma_w$?} \mc{addressed}
    of a stationary series with 
    marginal unit exponential distribution obtained by transforming an AR(1) process
    with $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ with series length 
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates 
    replicates of
    block bootstrap with $l = \lceil n^{1/3} \rceil$. 
    The error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:exp_sigma1}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_exp_phi_1}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for 
    the first-order autocorrelation coefficient of a stationary series
    with marginal unit exponential distribution obtained by transforming an AR(1) 
    process with 
    $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ with series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates 
    replicates of
    block bootstrap with $l = \lceil n^{1/3} \rceil$. 
    The error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:exp_phi1}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_exp_mu_2}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the marginal mean $\mu$ of a stationary series with marginal unit exponential
    distribution obtained by transforming an AR(1) process with
    $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ with series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates of
    block bootstrap with $l = \lceil 2n^{1/3} \rceil$. 
    The error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:exp_mu2}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_exp_sigma_2}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for
    the marginal standard deviation $\sigma_w$ %\eds{$\sigma_w$?} \mc{addressed}
    of a stationary series with 
    marginal unit exponential distribution obtained by transforming an AR(1) process
    with $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ with series length 
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates 
    replicates of
    block bootstrap with $l = \lceil 2n^{1/3} \rceil$. 
    The error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:exp_sigma2}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_exp_phi_2}
  \caption{Empirical coverage rates of different 95\% block bootstrap CIs for 
    the first-order autocorrelation coefficient of a stationary series
    with marginal unit exponential distribution obtained by transforming an AR(1) 
    process with 
    $\phi \in \{-0.4, 0.2, 0, 0.2, 0.4\}$ with series length
    $n \in \{100, 200, 400, 800, 1600, 3200\}$ based on 10,000 replicates 
    replicates of
    block bootstrap with $l = \lceil 2n^{1/3} \rceil$. 
    The error bars represent 95\% CIs of the real coverage rates.}
  \label{fig:exp_phi2}
\end{figure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/exp_phi_intervals}
  \caption{50 replicate percentile, BC, and recentered percentile CIs for the
    lag-1 autocorrelation coefficient 0.355 ($\phi = 0.4$)
    of a stationary series with marginal exponential distribution
    obtained by transforming an AR(1) process with $\phi = 0.4$ with
    sample size $n \in \{800, 1600\}$ ($l = \lceil n^{1/3} \rceil$). For 
    each replicate, the lower and upper bounds of the CIs are displayed, as well 
    as $\hat\theta_n$ (blue circle) and $\bar\theta_n^{B}$ (red cross).}
  \label{fig:eri}
\end{figure}

For an exponential marginal distribution, the empirical coverage rates for 
$\mu$, $\sigma_w$, %\eds{define $\sigma_w$} 
%\mc{added definitions for $\mu$ and $\sigma_w$ in the above paragraph} 
and 
the lag-1 autocorrelation coefficient, %\mc{replaced $\phi$ here, as I understand
%it, we are not using $\phi$ in reference to the temporal dependence of the 
%exponentially distributed series?}, 
as well as 95\%
confidence intervals of the real coverage are displayed in 
\textbf{Figures~\ref{fig:exp_mu1}--\ref{fig:exp_phi2}}. 
Additionally, a set of 50 CIs are displayed for each of the percentile, 
BC, and recentered percentile approaches for exponentially distributed
samples of $n \in \{800, 1600\}$ with lag-1 autocorrelation coefficient 0.355
($\phi = 0.4$) in Figure~\ref{fig:eri}.

It appears that a greater sample size is generally required for the bootstrap
CIs to cover the mean and standard deviation parameters in the exponential
margin case than in the normal margin case. However, the trends and patterns
discussed in the main text regarding the performance of various methods and
diverse parameters remain consistent. For example, Student's $t$ confidence
intervals still exhibit higher coverage rates in comparison to alternative
methods. Performance continues to be more favorable when temporal dependence is
negative rather than positive. Altering the block size results in the same 
changes in performance of different CIs. Of particular significance, the 
percentile, BC,
and BCA confidence intervals still display a decline in coverage accuracy for
the lag-1 autocorrelation coefficient as sample size increases as demonstrated
in Figure~\ref{fig:eri}. Both the percentile and BC intervals persist in
manifesting the same bias issue. On the other hand, the recentered percentile
confidence interval continues to be effective in estimating the temporal 
dependence %\mc{replaced $\phi$ again here} 
due to the
inherent unbiasedness of the original point estimator. In sum, the findings for
series that are marginally exponentially distributed closely mirror those
attained for series that are marginally normally distributed.

@@

We have included this comment in the Design subsection:

We only
used serial dependences as strong as 0.4, because we only seek to 
establish the 
general trend as the strength of the autocorrelation 
increases, and how it varies depending on the sign of the autocorrelation and 
the parameter of interest.
